---
title: "COVID-19-trends"
author: "p. polytes"
date: "9/16/2020"
output: html_document
---
```{r example code, echo=FALSE}
# read data
rm(list=ls())

library(utils)

#read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", 
                 na.strings = "", fileEncoding = "UTF-8-BOM")

# the code below shows how you may recode the country name's 
# first letter to a numeric alphabetic position.
data1 <- data[(data$dateRep=="11/08/2020"), ]  
CountryAb <- as.integer(as.factor(substr(data1$countriesAndTerritories,1,1)))

# build a linear model for the relationship between 
# Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 and 
# the alphabetic position of country name's first letter.
model1<- lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000~CountryAb, 
            data=data1)
summary(model1)
# On 8/11/2020, we did not find a significant relationship 
# between alphabetic order of country name and COVID-19 cases. 
# But what if we keep looking at other dates? 
```

Our goal will be to identify significant relationships between the COVID-19 cases and other variables contained in the ECDC data.

## Load packages

```{r Load packages}
library(tidyverse)
library(utils)
```


## Read data into R

```{r read data, echo=FALSE}
#read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv",
na.strings = "", fileEncoding = "UTF-8-BOM")
head(data)
```


## Explore relationship between number of characters in month name and COVID-19 cases

```{r month name}
#calculate number of characters in month name
month_name <- data %>%
  mutate(month_name = month.name[month]) %>%
  mutate(month_name_length = nchar(month_name)) %>%
  filter(month<=7)

#regress number of characters in month name on COVID-19 cases
model1<- lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000~month_name_length, data=month_name)
summary(model1)
```


## Explanation
The results above suggest that 1 unit increase in length of month name is associated with 5.2105 decrease in cumulative number of COVID-19 cases per 100000 population in the last 14 days. It is no doubt that there shouldn't be a causal relationship between two variables. We managed to achieve statistically significant p-value after limited sample to data in January to July. As we know, number of COVID-19 cases kept rising in those 7 months. Meanwhile, January and February happen to be the two longest names among the first 7 months. Thus, we leveraged this coincidence to manipulate the data and achieve statistically significant p-value. 

